New lower bounds for Schur and weak Schur numbers

### 2021-04-16 article/schur.tex
environnement proof -> \begin{proof}[\textsc{Proof.}]

\textsc{Proof :} The lower bound comes from Abbott and Hanson's construction. The upper bound comes from the 
fact that a SF-template with length \(p\) and \(n\) colors is also a partition of \([\![1, p]\!]\) into \(n\) sum-free subsets.

@@

We know proceed to prove the above mentioned theorem.\\

\textsc{Proof:} We denote by \(f\) the colouring associated to the \(SF\)-template of lenght \(p\)
and \(g\) the one associated to the sum-free partition of \([\![1,q]\!]\).

@@

\begin{itemize}
\item If \(u+v < p\), then \(x+y = (\alpha+\beta)p - (u+v) \) with \(u+v \in [\![0,p-1]\!] \).
We assume \(h(x+y) > k\) otherwise the expected result is trivial.
Since \(g(\alpha) = g(\beta)\), the sum-free condition garanties \(g(\alpha) \neq g(\alpha+\beta)\), 
thus \(h(x+y) = k + g(\alpha + \beta) \neq k + g(\alpha) = h(x)\).
\item If \(u+v \geqslant p\), then \(x+y = (\alpha+\beta-1)p - (u+v-p) \) with \(u+v-p \in [\![0,p-1]\!] \).
Because of the assumption \(h(x) > k\), we necessarily have \(f(p-u) = k+1 = f(p-v)\). 
Then \(f(p-(u+v-p)) = f(p-u + p-v) \leqslant k \) with the sum-free condition on \(f\), 
hence \(h(x+y) = f(p-(u+v-p)) \) by construction of \(h\). Thus \(h(x+y) \leqslant k < h(x)\).
\end{itemize}

\begin{remark}
The following proposition can help improve the additive constant of a SF-template. Although it does not allow us to 
improve the SF-templates we have found, the analogous of this proposition for WSF-templates (see next section) 
allows us to improve one of them.
\end{remark}

@@

\textsc{Proof :} The idea is the same as in the previous theorem. The only difference is the \(SF\) property inherited 
from the second \(SF\)-template.

>>>>

\begin{proof}[\textsc{Proof.}]
The lower bound comes from Abbott and Hanson's construction. The upper bound comes from the 
fact that a SF-template with length \(p\) and \(n\) colors is also a partition of \([\![1, p]\!]\) into \(n\) sum-free subsets.
\end{proof}

@@

We now proceed to prove the above mentioned theorem.\\

\begin{proof}[\textsc{Proof.}]
We denote by \(f\) the colouring associated to the \(SF\)-template of lenght \(p\)
and \(g\) the one associated to the sum-free partition of \([\![1,q]\!]\).

@@

\begin{itemize}
\item If \(u+v < p\), then \(x+y = (\alpha+\beta)p - (u+v) \) with \(u+v \in [\![0,p-1]\!] \).
We assume \(h(x+y) > k\) otherwise the expected result is trivial.
Since \(g(\alpha) = g(\beta)\), the sum-free condition garanties \(g(\alpha) \neq g(\alpha+\beta)\), 
thus \(h(x+y) = k + g(\alpha + \beta) \neq k + g(\alpha) = h(x)\).
\item If \(u+v \geqslant p\), then \(x+y = (\alpha+\beta-1)p - (u+v-p) \) with \(u+v-p \in [\![0,p-1]\!] \).
Because of the assumption \(h(x) > k\), we necessarily have \(f(p-u) = k+1 = f(p-v)\). 
Then \(f(p-(u+v-p)) = f(p-u + p-v) \leqslant k \) with the sum-free condition on \(f\), 
hence \(h(x+y) = f(p-(u+v-p)) \) by construction of \(h\). Thus \(h(x+y) \leqslant k < h(x)\).
\end{itemize}
\end{proof}

The following proposition can help improve the additive constant of a SF-template. Although it does not allow us to 
improve the SF-templates we have found, the analogous of this proposition for WSF-templates (see next section) 
allows us to improve one of them.

@@

\begin{proof}[\textsc{Proof.}]
The idea is the same as in the previous theorem. The only difference is the SF property inherited 
from the second SF-template.
\end{proof}

### 2021-04-16 article/weakschur.tex
amelioration introduction
environnement proof -> \begin{proof}[\textsc{Proof.}]

\section{Weak Schur numbers}

@@

\textsc{Proof :}Let  \(x\in [\![1,b]\!]\), let  \(y\in \mathbb{N}^*\) such that \(x+\pi(y)\leqslant a+b\)

@@

if \(x+\pi(y)\geqslant a\):
\begin{align*}
 \pi(x+y) & = (Id+a\mathds{1}_{ [\![0,b]\!]})(x+y \mod a)\\
& = (Id+a\mathds{1}_{ [\![0,b]\!]})(x+\pi(y)\mod a)\text{ since } \pi(y)=y \mod a \\
& = (Id+a\mathds{1}_{ [\![0,b]\!]})(x+\pi(y) - a) \\
& = x+\pi(y) - a +a\mathds{1}_{ [\![0,b]\!]})(x+\pi(y) - a) \\
& = x+\pi(y) - a +a \text{ \quad since } x+\pi(y) - a \in [\![0,b]\!] \\
& = x+\pi(y)
\end{align*}

@@

\textsc{Proof :}Let \((x,y)\in (\mathbb{N}^*)^2\),

\begin{align*}
 \\\pi(\pi(x)+\pi(y)) & = (Id+a\mathds{1}_{ [\![0,b]\!]})(\pi(x)+\pi(y)\text{ mod a})\\
& = (Id+a\mathds{1}_{ [\![0,b]\!]})((Id+a\mathds{1}_{ [\![0,b]\!]})(x \mod a)+(Id+a\mathds{1}_{ [\![0,b]\!]})(y \mod a)\ mod a) \\
& = (Id+a\mathds{1}_{ [\![0,b]\!]})(((x \mod a) + (y \mod a)) \mod a)\\
& =(Id+a\mathds{1}_{ [\![0,b]\!]})( x+y \mod a)\\
& =\pi(x+y)
\end{align*}

@@

\begin{proposition}
Let \((a,b)\in (\mathbb{N}^*)^2\), \(a>b\), let  \(x\in \mathbb{N}^*\), \(x=a\lambda(x)+\pi(x)-a\)
\end{proposition}
\textsc{Proof :}Let \((a,b)\in (\mathbb{N}^*)^2\), \(a>b\), let  \(x\in \mathbb{N}^*\),
\\\\ \(a\lambda(x)+\pi(x)-a=a\left\lfloor\dfrac{x-b-1}{a}\right\rfloor+(x \mod a)+\mathds{1}_{ [\![0,b]\!]}(x \mod a)\)

@@

if \(x \mod a \leqslant b\):
\(a\lambda(x)+\pi(x)-a=a \left( \left \lfloor \dfrac{x}{a} \right \rfloor - 1 \right)+\text{x mod a}+a=x\)

@@

\textsc{Proof :} The lower bound comes from the analogous of Abott and Hanson's construction for weak Schur numbers.
The upper bound comes from the fact that a WSF-template of length \(p\) with \(n\) colors is also a partition of
\([\![1, p]\!]\) into \(n\) sum-free subsets.

@@

\textsc{Proof:} Let \((q,n,b) \in (\mathbb{N}^*)^3\), let \( (p,k) \in (\mathbb{N}^*)^2\), let \(a=q-b\). \\
We denote by \(g\) the colouring associated to the partition of \([\![1,q]\!]\) and \(h\) the
one associated to the partition of \([\![1,p]\!]\).
\[ g : [\![1,q]\!] \longrightarrow [\![1,n]\!] \text{ and } (A_{g^{-1}(1)},...A_{g^{-1}(q)})\text{ is a \(b\)-WSF-template.}
\]
\[h : [\![1,p]\!] \longrightarrow [\![1,k]\!] \text{ and } \forall (x,y) \in [\![1,q]\!]^2 \text{, } h(x) = h(y)
\Longrightarrow h(x+y) \neq h(x)
\]

@@

\begin{itemize}
\item If \(\pi(x)+\pi(y)>q\), \(g(\pi(\pi(x)+\pi(y)) \neq g(\pi(x))\) (qv previous proposition)
\\Hence \(g(\pi(x+y))= g(\pi(\pi(x)+\pi(y))\neq g(\pi(x))\)
\\if \(g(\pi(x+y))=n, f(x+y)\geqslant n > f(x)\)
\\else, \(f(x+y)=g(\pi(x+y))\neq g(\pi(x))=f(x)\)
\item If  \(\pi(x)+\pi(y)\leqslant q\), \(g(\pi(\pi(x)+\pi(y)) \neq g(\pi(x))\) since \(g\) is sum-free for \(x>b\)
\\if \(g(\pi(x+y))=n, f(x+y)\geqslant n > f(x)\)
\\else, \(f(x+y)=g(\pi(x+y))=g(\pi(\pi(x)+\pi(y))\neq g(\pi(x))=f(x)\)

\end{itemize}

@@

\item If \(r(x)+r(y)>b\), \(\pi(x)+\pi(y)>2a+b\)
\\Since \(g\) is a \(b\)-WSF template, \(g(\pi(\pi(x)+\pi(y))) \neq n\)
\\Therefore, \(g(\pi(x+y)) \neq n\) ie \(x+y \in \mathcal{C}\)
\\Hence \(f(x+y) <n\leqslant f(x)\)
\item If \(r(x)+r(y) \leqslant b - a\), \(\pi(x)+\pi(y)\leqslant b+a\)
\\Since \(g\) is sum-free for \(x>b\), since \(g(\pi(x)) = g(\pi(y))=n\), \(g(\pi(x+y)) \neq g(\pi(x))=n\)
\\Hence, \(f(x+y) <n\leqslant f(x)\)
\end{itemize}

@@

\textsc{Proof :} The idea is the same as in the previous theorem. The only difference is the \(SF\) property inherited
from the second \(SF\)-template.

>>>>

\section{Weak Schur numbers}
\label{WeakSchur}

@@

\begin{proof}[\textsc{Proof.}]
Let  \(x\in [\![1,b]\!]\), let  \(y\in \mathbb{N}^*\) such that \(x+\pi(y)\leqslant a+b\)

@@

if \(x+\pi(y)\geqslant a\):
\begin{align*}
 \pi(x+y) & = (Id+a\mathds{1}_{ [\![0,b]\!]})(x+y \mod a)\\
& = (Id+a\mathds{1}_{ [\![0,b]\!]})(x+\pi(y)\mod a)\text{ since } \pi(y)=y \mod a \\
& = (Id+a\mathds{1}_{ [\![0,b]\!]})(x+\pi(y) - a) \\
& = x+\pi(y) - a +a\mathds{1}_{ [\![0,b]\!]})(x+\pi(y) - a) \\
& = x+\pi(y) - a +a \text{ \quad since } x+\pi(y) - a \in [\![0,b]\!] \\
& = x+\pi(y)
\end{align*}
\end{proof}

@@

\begin{proof}[\textsc{Proof.}]
Let \((x,y)\in (\mathbb{N}^*)^2\),

\begin{align*}
 \\\pi(\pi(x)+\pi(y)) & = (Id+a\mathds{1}_{ [\![0,b]\!]})(\pi(x)+\pi(y)\text{ mod a})\\
& = (Id+a\mathds{1}_{ [\![0,b]\!]})((Id+a\mathds{1}_{ [\![0,b]\!]})(x \mod a)+(Id+a\mathds{1}_{ [\![0,b]\!]})(y \mod a)\ mod a) \\
& = (Id+a\mathds{1}_{ [\![0,b]\!]})(((x \mod a) + (y \mod a)) \mod a)\\
& =(Id+a\mathds{1}_{ [\![0,b]\!]})( x+y \mod a)\\
& =\pi(x+y)
\end{align*}
\end{proof}

@@

\begin{proposition}
Let \((a,b)\in (\mathbb{N}^*)^2\), \(a>b\), let  \(x\in \mathbb{N}^*\), \(x=a\lambda(x)+\pi(x)-a\)
\end{proposition}
\begin{proof}[\textsc{Proof.}]
Let \((a,b)\in (\mathbb{N}^*)^2\), \(a>b\), let  \(x\in \mathbb{N}^*\),\\
\(a\lambda(x)+\pi(x)-a=a\left\lfloor\dfrac{x-b-1}{a}\right\rfloor+(x \mod a)+\mathds{1}_{ [\![0,b]\!]}(x \mod a)\)

@@

if \(x \mod a \leqslant b\):
\(a\lambda(x)+\pi(x)-a=a \left( \left \lfloor \dfrac{x}{a} \right \rfloor - 1 \right)+\text{x mod a}+a=x\)
\end{proof}

@@

\begin{proof}[\textsc{Proof.}]
The lower bound comes from the analogous of Abott and Hanson's construction for weak Schur numbers.
The upper bound comes from the fact that a WSF-template of length \(p\) with \(n\) colors is also a partition of
\([\![1, p]\!]\) into \(n\) sum-free subsets.
\end{proof}

@@

\begin{proof}[\textsc{Proof.}]
Let \((q,n,b) \in (\mathbb{N}^*)^3\), let \( (p,k) \in (\mathbb{N}^*)^2\), let \(a=q-b\). \\
We denote by \(g\) the colouring associated to the partition of \([\![1,q]\!]\) and \(h\) the
one associated to the partition of \([\![1,p]\!]\).
\[ g : [\![1,q]\!] \longrightarrow [\![1,n]\!] \text{ and } (A_{g^{-1}(1)},...A_{g^{-1}(q)})\text{ is a \(b\)-WSF-template.}
\]
\[h : [\![1,p]\!] \longrightarrow [\![1,k]\!] \text{ and } \forall (x,y) \in [\![1,q]\!]^2 \text{, } h(x) = h(y)
\Longrightarrow h(x+y) \neq h(x)
\]

@@

\begin{itemize}
\item If \(\pi(x)+\pi(y)>q\), \(g(\pi(\pi(x)+\pi(y)) \neq g(\pi(x))\) (qv previous proposition)
\\Hence \(g(\pi(x+y))= g(\pi(\pi(x)+\pi(y))\neq g(\pi(x))\)
\\if \(g(\pi(x+y))=n, f(x+y)\geqslant n > f(x)\)
\\else, \(f(x+y)=g(\pi(x+y))\neq g(\pi(x))=f(x)\)
\item If  \(\pi(x)+\pi(y)\leqslant q\), \(g(\pi(\pi(x)+\pi(y)) \neq g(\pi(x))\) since \(g\) is sum-free for \(x>b\)
\\if \(g(\pi(x+y))=n, f(x+y)\geqslant n > f(x)\)
\\else, \(f(x+y)=g(\pi(x+y))=g(\pi(\pi(x)+\pi(y))\neq g(\pi(x))=f(x)\)
\end{itemize}

@@

\item If \(r(x)+r(y)>b\), \(\pi(x)+\pi(y)>2a+b\)
\\Since \(g\) is a \(b\)-WSF template, \(g(\pi(\pi(x)+\pi(y))) \neq n\)
\\Therefore, \(g(\pi(x+y)) \neq n\) ie \(x+y \in \mathcal{C}\)
\\Hence \(f(x+y) <n\leqslant f(x)\)
\item If \(r(x)+r(y) \leqslant b - a\), \(\pi(x)+\pi(y)\leqslant b+a\)
\\Since \(g\) is sum-free for \(x>b\), since \(g(\pi(x)) = g(\pi(y))=n\), \(g(\pi(x+y)) \neq g(\pi(x))=n\)
\\Hence, \(f(x+y) <n\leqslant f(x)\)
\end{itemize}
\end{proof}

@@

\begin{proof}[\textsc{Proof.}]
The idea is the same as in the previous theorem. The only difference is the WSF property inherited
from the WSF-template.
\end{proof}
